Lenses With a Difference

### 2018-02-03 papers/DifferenceLenses/differencelenses.tex
Finished difference and lens proofs

% Macros
% Colors
\definecolor{dkblue}{rgb}{0,0.1,0.5}
\definecolor{dkgreen}{rgb}{0,0.6,0}
\definecolor{dkred}{rgb}{0.6,0,0}
\definecolor{dkpurple}{rgb}{0.7,0,0.4}
\definecolor{olive}{rgb}{0.4, 0.4, 0.0}
\definecolor{teal}{rgb}{0.0,0.5,0.5}
\definecolor{orange}{rgb}{0.9,0.6,0.2}
\definecolor{lightyellow}{RGB}{255, 255, 179}
\definecolor{lightgreen}{RGB}{170, 255, 220}
\definecolor{teal}{RGB}{141,211,199}
\definecolor{darkbrown}{RGB}{121,37,0}

@@

% % Author with single affiliation.
%\author{First1 Last1}
%\authornote{with author1 note}          %% \authornote is optional;
% % can be repeated if necessary
%\orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional \affiliation{
%\position{Position1}
%\department{Department1}              %% \department is recommended
%\institution{Institution1}            %% \institution is required
%\streetaddress{Street1 Address1}
%\city{City1}
%\state{State1}
%\postcode{Post-Code1}
%\country{Country1}                    %% \country is recommended }
%\email{first1.last1@inst1.edu}          %% \email is recommended

@@

One of these laws requires that the \pput component if a lens returns the
original source unmodified whenever the update to the view is a no-op.
Unfortunately, this (and other) laws do not specify the behaviour of the \pput
component when the update to the view is non-trivial. Consequently, a small
update to the view can result in a large update to the source. For example, if
the source and view are order agnostic, then a lens which updates records
positionally may be valid, even though such a lens may update all the records
in the source when only a single record is added at the beginning of the view.

This paper strengthens the traditional lens laws by adding a law that
constrains the behaviour of the \pput component of a lens even when the update
made to the view is non-trivial. To do this, we will first interprete sources
and views as {\em difference posets} which are posets where we can compute the
difference between any two elements of the poset. With this interpretation, we
will then express the new law which will require that the \pput component of a
lens preserve differences. For example, we will be able to invalidate
positional lenses defined on order agnostic data on the grounds that the \pput
component of positional lenses do not respect order agnostic differences.

@@

\section{Difference Posets}
\begin{definition}
A {\em difference poset (DP)} is a triple $(P, \rho, \delta)$ where $P$ is
a poset and $\rho : P \longrightarrow \mathbb{R}^{\geq 0}$ and $\delta : P
\times P \longrightarrow P$ are functions satisfying :
\begin{align}
&\delta(a, b) \leq a\tag{A1}\\
&\text{if } a < a',\text{ then } \rho(a) < \rho(a') \tag{A2}\\
&\text{if } a \leq a' \text{ then for all }b , \; \rho(\delta(b, a)) \geq
\rho(\delta(b, a'))\tag{A3}
\end{align}
We call $\rho$ the weight function of $(P, \rho, \delta)$ and $\delta$ its
difference function.
\end{definition}
\iffalse
Intuitively, $\pi_1(\delta(a, b))$ tells us how much $a$ differs from $a \wedge
b$ while $\pi_2(\delta(a, b))$ tells us how much $b$ differs from $a \wedge b$.
The advantage of the definition of difference functions is that $\delta(a, b)$
is defined even if $a \wedge b$ does not exist. Still it is instructive to
think of $a \wedge b$ to understand what the axioms are saying.

@@

The fourth law says that if $a \leq a'$ then the difference between $b$
and $a \wedge b$ is greater than or equal to the difference between $b$ and $a'
\wedge b$.
\fi
\begin{example}\label{embedding}
Let $P$ be any set with equality giving the poset structure. Given any
function $\rho: P \longrightarrow \mathbb{R}^{\geq 0}$ on $P$, and function
$\delta$ defined by $\delta(a, b) = a$, then $(P, \rho, \delta)$ is a difference
poset.
\end{example}
\begin{example}\label{naturalnumbers}
Let $P$ be the poset given by non-negative integers with the order $\leq$ and
weight function $\rho(a) = a$. Define $\delta$ by $\delta(a, b) = \max\{0,
a-b\}$. Then $(P, \rho, \delta)$ is a difference poset.
\end{example}
\begin{example}\label{divisibility}
Let $P$ be the poset given by non-negative integers with the division
relation and weight function $\rho(p_1^{a_1} \cdot \ldots \cdot p_n^{a_n}) =
a_1 + \ldots + a_n$. Define $\delta$ by $\delta(a, b) = a/gcd(a, b)$. Then $(P,
\rho, \delta)$ is a difference poset.
\end{example}
\begin{example}\label{boolean}
Let $X$ be a set and $P$ the poset given by the finite subsets of $2^X$ with the
containment relation and weight function $\rho(A) = |A|$. Define $\delta$
by $\delta(A, B) = A \setminus B$. Then $(P, \rho, \delta)$ is a difference
poset.
\end{example}
\begin{example}\label{general}
Generalizing \cref{divisibility} and \cref{boolean}, let $P$ be a complete
lattice and $\rho$ a strict order preserving map from $P$ to the non-negative
real numbers. Define $$\delta(a, b) = \rho\left(\bigwedge \{x \; | \; (a \wedge
b) \vee x = a\}\right)$$
Then $(P, \rho, \delta)$ is a difference poset.
\end{example}
\begin{proof}
Let $a, a', b \in P$.  Let $X = \{x \; | \; (a \wedge b) \vee x =  a\}$, $Y =
\{y \; | \; (b \wedge a) \vee y = b\}$ and $Y' = \{y' \; | \; (b \wedge a')
\vee y' = b\}$.
\begin{enumerate}
  \item[(A1)]
  Since $a \in X$ and $\delta(a, b) = \bigwedge X$, it follows that
  $\bigwedge X \leq a$.
  \item[(A2)]
  Follows from the fact that $\rho$ is a strict order preserving map from $P$ to
  the non-negative real numbers.
  \item[(A3)]
  Assume that $a \leq a'$. First we show that $Y \subseteq Y'$. Assume that
  $y \in Y$. Since $a \wedge b \leq a' \wedge b \leq b = (a \wedge b) \vee y$,
  it follows that, $(a' \wedge b) \vee y = (a \wedge b) \vee y = b$, hence $y
  \in Y'$. Consequently, $\bigwedge Y \geq \bigwedge Y'$ from which it follows
  that $$\rho(\delta(b, a)) =\rho\left(\bigwedge Y\right) \geq
  \rho\left(\bigwedge Y'\right) = \rho(\delta(b, a'))$$
  This completes the proof.
\end{enumerate}
\end{proof}
\iffalse
Now define a
function $d:
P \longrightarrow \mathbb{R}$ on $P$ by $$d(a, b) =
\begin{cases}
0 & \text{if } a = b\\
\rho(\pi_1(\delta(a, b))) + \rho(\pi_2(\delta(a, b))) & \text{otherwise}
\end{cases}$$
\begin{fact}
\begin{enumerate}
  \item
  $d(a,b) \geq 0$,
  \item
  $d(a, b) = 0$ if and only if $a = b$, and
  \item
  $d(a, b) = d(b, a)$
\end{enumerate}
\end{fact}
\begin{fact}
If $b \leq c$, then for all $b$, $d(a, b) \geq d(a, c)$. Thus if $b \leq c$,
$d(a, c) \leq d(a, b) + d(b, c)$
\end{fact}
\begin{question}
Under what conditions is $d$ a metric?
\end{question}
\fi
\section{New Difference Posets from Old}
Let $(P, \rho, \delta)$ and $(P', \rho', \delta)$ be difference posets.

\subsection{Product of Posets}
Define the product $P \times P'$ of $P$ and $P'$ to be the set $P \times P'$
with $\leq_{P \times P'} = \leq_P \times \leq_{P'}$. Define the weight function
$(\rho \times \rho')$ on $P \times P'$ by $(\rho \times \rho')(a, b) = \rho(a)
+ \rho(b)$. Define the difference function $\delta_P \times \delta_{P'}$ on $P
\times P$ by $$(\delta \times \delta')((a_1, b_1), (a_2, b_2)) = (\delta(a_1,
a_2), \delta(b_1, b_2))$$
\begin{claim}
$(P \times P', \rho \times \rho', \delta \times \delta)$ is a difference poset.
\end{claim}
\begin{proof}
We check that the definition that each of the axioms for a difference poset
holds.
\begin{enumerate}
  \item[(A1)]
  Since $\delta$ and $\delta'$ are difference functions, then $\delta(a_1, a_2)
  \leq a_1$ and $\delta'(b_1, b_2) \leq b_1$, hence $$
  (\delta \times \delta')((a_1, b_1), (a_2, b_2)) = (\delta(a_1, a_2),
  \delta(b_1, b_2)) \leq (a_1, b_1)$$
  \item[(A2)]
  Assume that $(a_1, b_1) < (a_2, b_2)$. Then since $\rho$ and $\rho'$ are
  weight functions, it follows that $\rho(a_1) < \rho(a_2)$ and
  $\rho'(b_1) < \rho'(b_2)$. Consequently,
  $$(\rho \times \rho')(a_1, b_1) = \rho(a_1) + \rho(b_1) < \rho(a_2) +
  \rho(b_2) = (\rho \times \rho)(a_2, b_2)$$
  \item[(A3)]
  Assume that $(a_1, b_1) \leq (a_2, b_2)$. Then since $\delta$ and $\delta'$ are
  difference functions, it follows that $\rho(\delta(a_3, b_1))
  \geq \rho(\delta(a_3, b_2))$ and $\rho'(\delta'(b_3, b_1)) \geq
  \rho'(\delta'(b_3, b_2))$.
  Consequently,
  \begin{align*}
((\rho \times \rho') \circ (\delta \times \delta'))((a_3, b_3), (a_1, b_1)))
&= (\rho \times \rho')(\delta(a_3, a_1), \delta'(b_3, b_1))\\
&= \rho(\delta(a_3, a_1)) + \rho'(\delta(b_3, b_1))\\
&\geq \rho(\delta(a_3, a_2)) + \rho'(\delta(b_3, b_2))\\
&= (\rho \times \rho')(\delta(a_3, a_2), \delta'(b_3, b_2))\\
&= ((\rho \times \rho') \circ (\delta \times \delta'))((a_3, b_3), (a_2,
b_2))\end{align*}

@@

a & \text{otherwise}

@@

  If $x, y \in P$ then $(\delta \sqcup \delta')(x, y) = \delta(x, y) \leq x$.
  A similar argument shows the same result if $x, y \in P'$. Otherwise $\delta
  \sqcup \delta'(a, b) = a$.

@@

  Assume that $a \leq_{P \sqcup P'} b$. If $a \leq_P b$,
  $$(\rho \sqcup \rho')(a) = \rho(a) < \rho(b) = (\rho \sqcup \rho')(b)$$
  If $a \leq_{P'} b$, then
  $$(\rho \sqcup \rho')(a) = \rho'(a) < \rho'(b) = (\rho \sqcup \rho')(b)$$

@@

  \in P$, then $\rho(\delta(c, a)) \geq \rho(\delta(c, a'))$ since $\delta$ is a
  difference function. If $c \in P'$, then $\delta(c, a) = \delta(c, b) = c$,
  hence $\rho(\delta(c, a)) \geq \rho(\delta(c, a))$. A similar argument shows
  that if $a \leq_{P'} b$, then for all $c \in P \sqcup P'$, $\rho(\delta(c, a))
  \geq \rho(\delta(c, a'))$. This completes the proof.

@@

$\Delta^*$ on $P^*$ by 
$$\Delta^*((x_1, \ldots, x_m), (y_1, \ldots, y_n)) = ((\delta(x_1, y_1), \ldots,
\delta(x_k, y_k), x_{k+1}, \ldots, x_m)$$

@@

$(P^*, \rho^*, \Delta^*)$ is a difference poset.

@@

  \leq x_i$ for $1 \leq i \leq \min\{m, n\}$, from which it follows that
  $\Delta^*((x_1, \ldots, x_m), (y_1, \ldots, y_n)) \leq (x_1, \ldots, x_m)$

@@

  Assume that $(x_1, \ldots, x_m) < (y_1, \ldots, y_n)$. Then $m \leq
  n$ and $\rho(x_k) < \rho(y_k)$ for some $1 \leq k \leq m$ and $\rho(x_i)
  \leq \rho(y_i)$ for all $1 \leq i \leq m$, from which it follows that
  $$\rho^*(x_1, \ldots, x_m) = \sum_{i=1}^m \rho(x_i) < \sum_{i=1}^{n}\rho(y_i)
  = \rho^*(y_1, \ldots, y_n)$$

@@

  \min\{m', n\}$. By the definition of $\Delta^*$,

@@

\Delta^*((z_1, \ldots, z_n), (x_1, \ldots, x_m)) &= (\delta(z_1, x_1),
\delta(z_k, x_k), z_{k+1}, \ldots, z_n)\\
\Delta^*((z_1, \ldots, z_n), (y_1, \ldots, y_{m'})) &= (\delta(z_1, y_1),
\delta(z_k, y_{k'}), z_{k'+1}, \ldots, z_n)

@@

$$\Delta^*((z_1, \ldots, z_n), (x_1, \ldots, x_m)) \geq \Delta^*((z_1,

@@

Now let $x = (x_1, \ldots, x_m), \; y = (y_1, \ldots, y_n)$. We define the
function $\Gamma^*$ on $P^*$ by taking the difference of $x_i$ and
$y_{\varphi(i)}$ for some ``maximal'' increasing injective function $\varphi :
A \longrightarrow [n]$ where $A$ is a subset of $[m]$, except that we eliminate
all $x_i$ where $x_i = y_{\varphi(i)}$.

For example, assume that $P$ is the set consisting of the letters of the
alphabet where the poset structure is given by equality, with $\rho(a) = 1$
and $\delta(a, b) = a$ for all $a, b \in P$. Then $\Gamma^*((a, c, d, f, b, e),
(a, b, d, e)) = (c, f, b)$ where we eliminate $a$, $d$ and $e$ from the first
sequence. However, even though $b$ occurs in both sequences, we are not allowed
to eliminate both $b$ and $d$ since these two characters occur in the order $d,
b$ in the first sequence and in the order $b, d$ in the second sequence. If we
were to eliminate both of them, then the matching would have to have a
crossing, thus contradicting the monotonicity of $\varphi$.

Given any increasing injective $\varphi : A \longrightarrow [n]$ for some
subset $A$ of $[m]$, define $\Gamma^*_{\varphi}(x, y)$ to be the difference
between $x$ and $y$ computed using $\varphi$. Then the maximal $\varphi$ is the
$\varphi$ that minimizes $\rho^*(\Gamma^*_{\varphi}(x, y))$. If there are
multiple such $\varphi$, we choose the one that maximizes $|A|$. If there are
still multiple such $\varphi$, we choose the one for which $A$ is
lexicographically minimal.
\begin{claim}\label{nocrossingproof}
$(P^*, \rho^*, \Gamma^*)$ is a difference poset.
\end{claim}
\begin{proof}
Let $x = (x_1, \ldots, x_m)$, $y = (y_1, \ldots, y_n)$ and $z = (z_1, \ldots,
z_k)$. We check that each of the axioms for a difference poset holds.
\begin{enumerate}
  \item[(A1)]
  Since each element of $\Gamma^*(x, y)$ is either equal to $\Gamma(x_i,
  y_{\varphi(i)})$ for some $\varphi$, or to $x_j$ for some $j$, and since these
  elements occur in increasing order, it follows that $\Gamma^*(x, y) \leq x$.
  \item[(A2)]
  Assume that $(x_1, \ldots, x_m) < (y_1, \ldots, y_n)$. Then $m \leq
  n$ and there exists an increasing injective function $\varphi:[m]
  \longrightarrow [n]$ such that $\rho(x_k) < \rho(y_{\varphi(k)})$ for some $1
  \leq k \leq m$ and $\rho(x_i) \leq \rho(y_{\varphi(i)})$ for all $1 \leq i
  \leq m$. It thus follows that $$\rho^*(x_1, \ldots, x_m) = \sum_{i=1}^m
  \rho(x_i) < \sum_{i=1}^{n}\rho(y_i) = \rho^*(y_1, \ldots, y_n)$$
  \item[(A3)]
  Assume that $(x_1, \ldots, x_m) \leq (y_1, \ldots, y_n)$. Assume that
  $\Gamma^*(z, x) = \Gamma^*_{\varphi}(z, x)$ for some $\varphi$. Then
  $\rho^*(\Gamma^*_{\varphi} (z, x)) \geq \rho^*(\Gamma^*_{\varphi}(z, y))$
  since $\Gamma$ is a difference function by assumption. Consequently
  $\rho^*(\Gamma^*_{\varphi} (z, x)) \geq \rho^*(\Gamma^*_{\varphi}(z, y)) \geq
  \rho^*(\Gamma^*(z, y))$ since $\Gamma^*(z, y)$ is computed using a $\psi$ that
  minimizes the quantity $\rho^*(\Gamma^*_{\psi}(z, y))$. This completes the
  proof.
\end{enumerate}
\end{proof}
\subsubsection{Quotienting $P^*$ to Ignore Order}
We may try to define an even finer poset structure on $P^*$ by $x \leq y$ if and
only if $x = (x_1, \ldots, x_m), \; y = (y_1, \ldots, y_n)$, $m \leq n$ and
there exists an injective function $\varphi : [m] \longrightarrow [n]$ such
that $x_i \leq y_{\varphi(i)}$ for $1 \leq i \leq m$: we drop the requirement
that $\varphi$ be increasing.

@@

Recall that we wished to define an even finer poset structure on $P^*$ by $x
\leq y$ if and only if $x = (x_1, \ldots, x_m), \; y = (y_1, \ldots, y_n)$, $m
\leq n$ and there exists an injective function $\varphi : [m] \longrightarrow
[n]$ such that $x_i \leq y_{\varphi(i)}$ for $1 \leq i \leq m$. We observed
that with this finer poset structure on $P^*$, we need to identify
$(x_{\varphi(1)}, \ldots, x_{\varphi(n)})$ and $(x_{\psi(1)}, \ldots,
x_{\psi(n)})$ for all permutations $\varphi$ and $\psi$. In other words we take
the poset structure $\leq_{NC}$ (NC for non-crossing) described in
\cref{nocrossing} and quotient $P^*$ by the equivalence relation which
identifies $(x_{\varphi(1)}, \ldots, x_{\varphi(n)})$ and $(x_{\psi(1)},
\ldots, x_{\psi(n)})$ for all permutations $\varphi$ and $\psi$

@@

\subsubsection{Default}
Let $\ell : S \longrightarrow V$ be a lens. Let $f : V \longrightarrow S$ be any

>>>>

% Macros Colors
\definecolor{dkblue}{rgb}{0,0.1,0.5}
\definecolor{dkgreen}{rgb}{0,0.6,0}
\definecolor{dkred}{rgb}{0.6,0,0}
\definecolor{dkpurple}{rgb}{0.7,0,0.4}
\definecolor{olive}{rgb}{0.4, 0.4, 0.0}
\definecolor{teal}{rgb}{0.0,0.5,0.5}
\definecolor{orange}{rgb}{0.9,0.6,0.2}
\definecolor{lightyellow}{RGB}{255, 255, 179}
\definecolor{lightgreen}{RGB}{170, 255, 220}
\definecolor{teal}{RGB}{141,211,199}
\definecolor{darkbrown}{RGB}{121,37,0}

@@

% % Author with single affiliation.
% \author{First1 Last1} \authornote{with author1 note}          %% \authornote
% is optional; % can be repeated if necessary \orcid{nnnn-nnnn-nnnn-nnnn}       
%      %% \orcid is optional \affiliation{ \position{Position1}
% \department{Department1}              %% \department is recommended
% \institution{Institution1}            %% \institution is required
% \streetaddress{Street1 Address1} \city{City1} \state{State1}
% \postcode{Post-Code1} \country{Country1}                    %% \country is
% recommended } \email{first1.last1@inst1.edu}          %% \email is recommended

@@

One of these laws requires that the \pput component of a lens returns the
original source unmodified whenever the update to the view is a no-op.
Unfortunately, this law does not specify the behaviour of the \pput
component when the update to the view is non-trivial. Consequently, a small
update to the view can result in a large update to the source. For example, if
the source and view are order agnostic, then a lens which updates records
positionally may be valid, even though such a lens may update all the records
in the source when only a single record is added at the beginning of the view.

This paper strengthens the traditional lens laws by adding a law that
constrains the behaviour of the \pput component of a lens even when the update
made to the view is non-trivial. To do this, we will first interprete sources
and views as {\em difference posets} which are posets where we can measure the
difference between any two elements of the poset. We will then use this
difference structure on sources and views to express the new lens law which will
require that the \pput component of a lens respect differences. For example, we
will be able to invalidate positional lenses defined on order agnostic data on
the grounds that the \pput component of a positional lenses does not respect
order agnostic differences.

@@

\section{Difference Posets}
\begin{definition}
A {\em difference poset (DP)} is a triple $(P, \rho, \delta)$ where $(P, \rho)$
is a graded poset $\delta : P \times P \longrightarrow \mathbb{N}$ is a function
satisfying :
\begin{align}
&\delta(a, b) \leq \rho(a)\tag{A1}\\
&\text{if } a \leq a',\text{ then } \delta(a, b) = 0\tag{A2}\\
&\text{if } a \leq a' \text{ then for all }b , \; \delta(b, a) \geq
\delta(b, a')\tag{A3}
\end{align}
We call $\rho$ the weight function of $(P, \rho, \delta)$ and $\delta$ its
difference function.
\end{definition}
\iffalse
Intuitively, $\pi_1(\delta(a, b))$ tells us how much $a$ differs from $a \wedge
b$ while $\pi_2(\delta(a, b))$ tells us how much $b$ differs from $a \wedge b$.
The advantage of the definition of difference functions is that $\delta(a, b)$
is defined even if $a \wedge b$ does not exist. Still it is instructive to
think of $a \wedge b$ to understand what the axioms are saying.

@@

The fourth law says that if $a \leq a'$ then the difference between $b$
and $a \wedge b$ is greater than or equal to the difference between $b$ and $a'
\wedge b$.
\fi
\begin{example}\label{embedding}
Let $P$ be any set with equality giving the poset structure. Define
$\rho(a) = 1$ for all $a$ and $\delta(a, b) = \delta_{a, b}$ where
$$\delta_{a, b} =
\begin{cases}
0 & \text{if } a = b\\
1 & \text{otherwise}
\end{cases}$$
\end{example}
Then $(P, \rho, \delta)$ is a difference poset.
\begin{example}\label{naturalnumbers}
Let $P$ be the poset given by non-negative integers with the order $\leq$ and
weight function $\rho(a) = a$. Define $\delta$ by $\delta(a, b) = \max\{0,
a-b\}$. Then $(P, \rho, \delta)$ is a difference poset.
\end{example}
\begin{example}\label{divisibility}
Let $P$ be the poset given by non-negative integers with the division
relation and weight function $\rho(p_1^{a_1} \cdot \ldots \cdot p_n^{a_n}) =
a_1 + \ldots + a_n$. Define $\delta$ by $\delta(a, b) = \rho(a/gcd(a, b))$. Then
$(P, \rho, \delta)$ is a difference poset.
\end{example}
\begin{example}\label{boolean}
Let $X$ be a set and $P$ the poset given by the finite subsets of $2^X$ with the
containment relation and weight function $\rho(A) = |A|$. Define $\delta$
by $\delta(A, B) = \rho(A \setminus B)$. Then $(P, \rho, \delta)$ is a
difference poset.
\end{example}
\begin{example}\label{general}
Generalizing \cref{divisibility} and \cref{boolean}, let $P$ be a complete
lattice and $\rho$ a strict order preserving map from $P$ to the non-negative
real numbers such that $\rho(a) = 0$ if and only if $a = \bigwedge P$. Define
$$\delta(a, b) = \rho\left(\bigwedge \{x \; | \; (a \wedge b) \vee x =
a\}\right)$$ Then $(P, \rho, \delta)$ is a difference poset.
\end{example}
\begin{proof}
Let $a, a', b \in P$.  Let $X = \{x \; | \; (a \wedge b) \vee x =  a\}$, $Y =
\{y \; | \; (b \wedge a) \vee y = b\}$ and $Y' = \{y' \; | \; (b \wedge a')
\vee y' = b\}$.
\begin{enumerate}
  \item[(A1)]
  Since $a \in X$ and $\delta(a, b) = \bigwedge X$, it follows that
  $\bigwedge X \leq a$.
  \item[(A2)]
  If $a \leq b$, then $(a \wedge b) = a$ hence $\bigwedge X = \bigwedge P$.
  Consequently, $\delta(a, a') = \left(\bigwedge X\right) =
  \left(\bigwedge P \right) = 0$
  \item[(A3)]
  Assume that $a \leq a'$. First we show that $Y \subseteq Y'$. Assume that
  $y \in Y$. Since $a \wedge b \leq a' \wedge b \leq b = (a \wedge b) \vee y$,
  it follows that, $(a' \wedge b) \vee y = (a \wedge b) \vee y = b$, hence $y
  \in Y'$. Consequently, $\bigwedge Y \geq \bigwedge Y'$ from which it follows
  that $$\delta(b, a) =\rho\left(\bigwedge Y\right) \geq
  \rho\left(\bigwedge Y'\right) = \delta(b, a')$$
  This completes the proof.
\end{enumerate}
\end{proof}

\section{New Difference Posets from Old}
Let $(P, \rho, \delta)$ and $(P', \rho', \delta)$ be difference posets.

\subsection{Product of Posets}
Define the product $P \times P'$ of $P$ and $P'$ to be the set $P \times P'$
with $\leq_{P \times P'} = \leq_P \times \leq_{P'}$. Define the weight function
$(\rho \times \rho')$ on $P \times P'$ by $(\rho \times \rho')(a, b) = \rho(a)
+ \rho(b)$. Define the difference function $\delta_P \times \delta_{P'}$ on $P
\times P$ by $$(\delta \times \delta')((a_1, b_1), (a_2, b_2)) = \delta(a_1,
a_2) + \delta(b_1, b_2)$$
\begin{claim}
$(P \times P', \rho \times \rho', \delta \times \delta)$ is a difference poset.
\end{claim}
\begin{proof}
We check that the definition that each of the axioms for a difference poset
holds.
\begin{enumerate}
  \item[(A1)]
  Since $\delta$ and $\delta'$ are difference functions, then $\delta(a_1, a_2)
  \leq \rho(a_1)$ and $\delta'(b_1, b_2) \leq \rho(b_1)$, hence $$
  (\delta \times \delta')((a_1, b_1), (a_2, b_2)) = \delta(a_1, a_2) +
  \delta(b_1, b_2) \leq \rho(a_1) + \rho(b_1) = \rho(a_1, b_1)$$
  \item[(A2)]
  Assume that $(a_1, b_1) \leq (a_2, b_2)$. Then $a_1 \leq a_2$ and $b_1 \leq
  b_2$. Since $\delta$ is a difference function, $\delta(a_1, a_2) = 0$ and
  $\delta(b_1, b_2) = 0$, hence
  $$\delta((a_1, b_1), (a_2, b_2)) = \delta(a_1, a_2) + \delta(b_1, b_2) = 0$$
  \item[(A3)]
  Assume that $(a_1, b_1) \leq (a_2, b_2)$. Then since $\delta$ and $\delta'$
  are difference functions, it follows that $\delta(a_3, b_1)
  \geq \delta(a_3, b_2)$ and $\delta'(b_3, b_1) \geq \delta'(b_3, b_2)$.
  Consequently,
  \begin{align*}
(\delta \times \delta')((a_3, b_3), (a_1, b_1)))
&=\delta(a_3, a_1) + \delta'(b_3, b_1)\\
&\geq \delta(a_3, a_2) + \delta(b_3, b_2)\\
&= \delta(a_3, a_2) + \delta'(b_3, b_2)\\
&= (\delta \times \delta')((a_3, b_3), (a_2, b_2))
\end{align*}

@@

\rho(a) & \text{otherwise}

@@

  If $x, y \in P$ then $(\delta \sqcup \delta')(x, y) = \delta(x, y) \leq
  \rho(x) = (\rho \sqcup \rho')(y)$. A similar argument shows the same result if
  $x, y \in P'$. Otherwise $\delta \sqcup \delta'(a, b) = \rho(a) = (\rho \sqcup
  \rho')(y)$.

@@

  Assume that $a \leq_{P \sqcup P'} b$. If $a \leq_P b$, then $\delta(a, b) =
  \delta_P(a, b) = 0$ and if $a \leq_{P'} b$, then $\delta(a, b) =
  \delta_{P'}(a, b) = 0$ since $\delta_P$ and $\delta_{P'}$ are difference

@@

  \in P$, then $\delta(c, a) \geq \delta(c, b)$ since $\delta_P$ is a
  difference function. If $c \in P'$, then $\delta(c, a) = \delta(c, b) =
  \rho(c)$. A similar argument shows that if $a \leq_{P'} b$, then for all $c
  \in P \sqcup P'$, $\delta(c, a) \geq \delta(c, a')$. This completes the
  proof.

@@

$\delta^*$ on $P^*$ by
$$\delta^*((x_1, \ldots, x_m), (y_1, \ldots, y_n)) =
\sum_{i=1}^{\min\{m, n\}} \delta(x_i, y_i) + \sum_{j > \min\{m, n\}}^n \rho(x_j)

@@

$(P^*, \rho^*, \delta^*)$ is a difference poset.

@@

  \leq \rho(x_i)$ for $1 \leq i \leq \min\{m, n\}$, from which it follows that
\delta^*((x_1, \ldots, x_m), (y_1, \ldots, y_n)) &=
\sum_{i=1}^{\min\{m, n\}} \delta(x_i, y_i) + \sum_{j > \min\{m, n\}}^n
\rho(x_j)\\
&\leq  \sum_{i=1}^{\min\{m, n\}} \rho(x_i) + \sum_{j > \min\{m, n\}}^n
\rho(x_j)\\
&= \rho(x_1, \ldots, x_n)

@@

Assume that $(x_1, \ldots, x_m) \leq (y_1, \ldots, y_n)$. Then $m \leq
n$ and since $\delta$ is a difference function, then $\delta(x_i, y_i) = 0$
for $1 \leq i \leq m$, from which it follows that
\delta^*((x_1, \ldots, x_m), (y_1, \ldots, y_n)) = \sum_{i=1}^m \delta(x_i,
y_i) = \sum_{i=1}^m 0 = 0

@@

\min\{m', n\}$. By the definition of $\delta_{NC}^*$,

@@

\delta^*((z_1, \ldots, z_n), (x_1, \ldots, x_m)) &= \sum_{i=1}^{k} \delta(x_i,
y_i) + \sum_{j > \min\{m, n\}}^n \rho(x_j)\\
\delta^*((z_1, \ldots, z_n), (y_1, \ldots, y_{m'})) &= \sum_{i=1}^{k'}
\delta(x_i, y_i) + \sum_{j > \min\{m, n\}}^n \rho(x_j)

@@

$$\delta^*((z_1, \ldots, z_n), (x_1, \ldots, x_m)) \geq \delta^*((z_1,

@@

Now let $x = (x_1, \ldots, x_m), \; y = (y_1, \ldots, y_n)$. We define the
function $\delta_{NC}^*$ on $P^*$ by taking the difference of $x_i$ and
$y_{\varphi(i)}$ for some ``maximal'' increasing injective function $\varphi :
A \longrightarrow [n]$ where $A$ is a subset of $[m]$ that minimizes $\sum_{i \in
A }\delta(x_i, y_{\varphi(i)}) + \sum_{j \not \in A}\rho(x_j)$. If there are
multiple such $\varphi$, we choose a canonical one.

For a maximal $\varphi$, we define $\delta^*_{NC}$ by
$$\delta^*_{NC}((x_1, \ldots, x_m), (y_1, \ldots, y_n)) = \sum_{i \in
A }\delta(x_i, y_{\varphi(i)}) + \sum_{j \not \in A}\rho(x_j)$$
For example, assume that $P$ is the set consisting of the letters of the
alphabet where the poset structure is given by equality, with $\rho(a) = 1$
and $\delta(a, b) = \begin{cases}0 & \text{if } a = b\\ 1 & 
\text{otherwise}\end{cases}$ for all $a, b \in P$. Then $\delta_{NC}^*((a, c,
d, f, b, e), (a, b, d, e)) = 3$ where we match $a$, $d$ and $e$ between the two
sequences thus leave $c, b$ and $f$ intact. However, even though $b$ occurs in
both sequences, we are not allowed to match $b$ and $d$ since these two
characters occur in the order $d, b$ in the first sequence and in the order $b,
d$ in the second sequence. If we were to eliminate both of them, then the
matching would have to have a crossing, thus contradicting the monotonicity of
$\varphi$.
\begin{claim}\label{nocrossingproof}
$(P^*, \rho^*, \delta_{NC}^*)$ is a difference poset.
\end{claim}
\begin{proof}
Let $x = (x_1, \ldots, x_m)$, $y = (y_1, \ldots, y_n)$ and $z = (z_1, \ldots,
z_k)$. We check that each of the axioms for a difference poset holds.
\begin{enumerate}
  \item[(A1)]
  Since each term in $\delta_{NC}^*(x, y)$ is either equal to $\delta(x_i,
  y_{\varphi(i)})$ for some $\varphi$, or to $\rho(x_j)$ for some $j$, it
  follows that $\delta_{NC}^*(x, y) \leq \rho(x)$.
  \item[(A2)]
  Assume that $(x_1, \ldots, x_m) \leq (y_1, \ldots, y_n)$. Then $m \leq
  n$ and there exists an increasing injective function $\varphi:[m]
  \longrightarrow [n]$ such that $x_k \leq y_{\varphi(k)}$ for all $1
  \leq k \leq m$. Consequently, $\delta^*_{NC, \varphi}((x_1, \ldots, x_m),
  (y_1, \ldots, y_m)) = 0$, hence $\delta^*_{NC}((x_1, \ldots, x_m), (y_1,
  \ldots, y_m)) = 0$
  \item[(A3)]
  Assume that $(x_1, \ldots, x_m) \leq (y_1, \ldots, y_n)$. Assume that
  $\delta_{NC}^*(z, x) = \delta_{NC, \varphi}^*(z, x)$ for some
  $\varphi$.
  Then $\delta_{NC, \varphi}^* (z, x) \geq \delta_{NC, \varphi}^*(z, y)$ since
  $\delta$ is a differene function. Consequently $\delta_{NC, \varphi}^* (z, x)
  \geq \delta_{NC, \varphi}^*(z, y) \geq \delta_{NC, \varphi}^*(z, y)$ since
  $\delta_{NC}^*(z, y)$ is computed using a $\psi$ that minimizes the quantity
  $\delta_{NC, \psi}^*(z, y)$. This completes the proof.
\end{enumerate}
\end{proof}
\subsubsection{Quotienting $P^*$ to Ignore Order}
We may try to define an even finer poset structure on $P^*$ by $x \leq y$ if and
only if $x = (x_1, \ldots, x_m), \; y = (y_1, \ldots, y_n)$, $m \leq n$ and
there exists an injective function $\varphi : [m] \longrightarrow [n]$ such
that $x_i \leq y_{\varphi(i)}$ for $1 \leq i \leq m$: we drop the requirement
that $\varphi$ be increasing.

@@

Recall that we wished to define an even finer poset structure on $P^*$ by $x
\leq y$ if and only if $x = (x_1, \ldots, x_m), \; y = (y_1, \ldots, y_n)$, $m
\leq n$ and there exists an injective function $\varphi : [m] \longrightarrow
[n]$ such that $x_i \leq y_{\varphi(i)}$ for $1 \leq i \leq m$. We observed
that with this finer poset structure on $P^*$, we need to identify
$(x_{\varphi(1)}, \ldots, x_{\varphi(n)})$ and $(x_{\psi(1)}, \ldots,
x_{\psi(n)})$ for all permutations $\varphi$ and $\psi$. In other words we take
the poset structure $\leq_{NC}$ described in \cref{nocrossing} and take
the quotient of $P^*$ by the equivalence relation which identifies
$(x_{\varphi(1)}, \ldots, x_{\varphi(n)})$ and $(x_{\psi(1)}, \ldots,
x_{\psi(n)})$ for all permutations $\varphi$ and $\psi$

@@

\subsubsection{Positional Iteration}
Let $\ell: S \Leftrightarrow V$ be a lens. Assume that $\rho(\ell.\create \;
